### 2015-11-13 test.tex
Abstract added

\documentclass[a4paper,man,natbib]{apa6}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}

\title{A Review of the Movie \textit{Apocalypse Now}}
\shorttitle{A Review of the Movie \textit{Apocalypse Now}}
\author{Samuel Pilla}
\affiliation{Missouri University of Science and Technology}

\abstract{This paper will explore the applications of Abnormal Psychology in the film \textit{Apocalypse Now}, including correct and incorrect aspects of symptoms, diagnosis, and other disorders that are depicted in the movie. Using the DSM-V as the source for determining the accuracy of disorders, the characters in the film, primarily Captain Benjamin L. Willard and Colonel Walter E. Kurtz, will be thoroughly examined to outline which disorder, or disorders, they exhibit and which the film may inaccurately depict. Because the film is based in the Vietnam War, the perspective when analyzing the characters will be shifted to fit the setting, thus potentially causing a diagnosis to be different than the diagnosis of a “normal” life experience.}

\begin{document}
\maketitle

\section{Introduction}

Your introduction goes here! Some examples of commonly used commands and features are listed below, to help you get started.

If you have a question, please use the support box in the bottom right of the screen to get in touch.

\section{SomeApplication to Abnormal Psychology}
\label{sec:examples}

\subsection{Aspects of the Movie that were Correct}

Use section and subsection commands to organize your document. \LaTeX{} handles all the formatting and numbering automatically. Use ref and label commands for cross-references.

\subsection{Aspects of the Movie that were Incorrect}

You can add inline TODO comments with the todonotes package, like this:
\todo[inline, color=green!40]{This is an inline comment.}

\subsection{Conclusion}

LaTeX automatically generates a bibliography in the APA style from your .bib file. The citep command generates a formatted citation in parentheses \citep{Lamport1986}. The cite command generates one without parentheses. LaTeX was first discovered by \cite{Lamport1986}.

\pagebreak

\subsection{References}
Coppola, F. F. (Producer), \& Coppola, F. F. (Director). (1979). \textit{Apocalypse Now} [Motion picture]. United States: Omni Zoetrope Studios.

American Psychiatric Association. (2013). Diagnostic and statistical manual of mental disorders (5th ed.). Washington, DC: Author.

\end{document}

%
% Please see the package documentation for more information
% on the APA6 document class:
%
% http://www.ctan.org/pkg/apa6
%